

# Notes on taxize integration

```{r}
library('taxize')
```


## General thoughts

* Add a `simplify` option for all functions that whose focus is to return information that could be encoded in `taxon_name`, `taxon_id`, `taxon`, or `taxon_rank` vectors, but all return other associated information in tables. If `simplify = TRUE` the just the vector is returned, if `simplify = FALSE` then the vector is the first column in a table containing the other information.


## gnr_resolve

```{r}
temp <- gnr_resolve(names = c("Helianthos annus", "Homo saapiens"))
head(temp)
```

The `matched_name` could be a `taxon_name` or `taxon` vector and the `data_source_title` could be encoded in its `taxon_id` field.


## get_* functions

```{r}
mynames <- c("Helianthus annuus ssp. jaegeri", "Helianthus annuus ssp. lenticularis", "Helianthus annuus ssp. texanus")
(tsn <- get_tsn(mynames, accepted = FALSE))
```

These could return `taxon_id` vectors, which would store the database in their `taxon_db` fields instead of using a class for the database.
The other attributes could remain the same or be put returned as columns in a table.

  

## accepted names functions

```{r}
itis_acceptname(searchtsn = tsn)
```

The `acceptedname`, `acceptedtsn`, and `author` columns could be converted to a `taxon_name` vector column.
Alternativly, a vector of `taxon` objects could be returned with the accepted as the first synonym and the query as the second.

## classification

```{r}
specieslist <- c("Abies procera","Pinus contorta")
classification(specieslist, db = 'itis')
```

This could return a `classification` vector.


## as.*id functions

Would probably be replaced by the `taxon_id` constructor to replicate the `check = FALSE` behavior.
For the `check = TRUE` behavior, perhaps `taxon_id` class should have a `validated` boolean field and then these functions could be replaced with a `check_taxon_id` function that populates that field.
This would allow a vector of IDs from different databases to be validated.


## downstream


```{r}
downstream("Dasypodidae", downto = "Species", db = "col")
downstream("Dasypodidae", downto = "Species", db = "col", intermediate = TRUE)
```

When `intermediate = FALSE`, the return value could be a list of `taxon_name` vectors or the columns with ids, names, ranks, and name status could be replaved with a `taxon_name` vector.

When `intermediate = TRUE`, the same as `intermediate = FALSE` as described above could be used, or a list of `taxonomy` could be returned.


## children

```{r}
children("Pinaceae", db = "ncbi")
```

Can also be list of `taxon_name` or tables with `taxon_name` as a column.


## genbank2uid


```{r}
genbank2uid(id = 'AJ748748')
```

Can return `taxon_id` vector.


## apgOrders and apgFamilies

```{r}
x = apgOrders()
head(x)
```

Can return a `taxon` vector so synonyms can be included.


## apg_lookup

```{r}
apg_lookup(taxa = "Hyacinthaceae", rank = "family")
apg_lookup(taxa = "Asteraceae", rank = "family")
```


Can return a `taxon_name` with the the db set to `apg`


## as.data.frame.*id

```{r}
as.data.frame(as.boldid(c("1973","101009","98597"), check=FALSE))
```

This functionality could mostly be handled by `as.data.frame.taxon_id`, except for the `multiple_matches` `pattern_match` columns.
Perhaps these functions can call  `as.data.frame.taxon_id` and add on those columns


## bold_search

```{r}
bold_search(name=c("Apis","Puma concolor"))
bold_search(id=c(88899, 88899), dataTypes="all")
bold_search(id=c(88899, 88899), dataTypes="stats")
bold_search(id=c(88899, 88899), dataTypes="geo")
```

Can replace `taxid`, `taxon`, and  `tax_rank` columns with a `taxon_name` column


## class2tree

Perhaps this could be modified to take `taxonomy` class as input, since that could be the output of `classifications`.
If so, it might make more sense to be in the `taxa` package, or another focused on visluizing taxonomic info.


## comm2sci

```{r}
comm2sci(commnames='american black bear')
comm2sci(commnames='american black bear')
comm2sci(commnames='black bear', db='itis', simplify = FALSE)
```

Could return a list of `taxon_name` vectors


## eol_search

```{r}
x <- eol_search(terms='Homo')
x
```

pageid, name, link columns could be replaced by `taxon_name` vector.


## eubon_search

```{r}
eubon_search("Prionus")
```

A lot of these columns could be made into a `taxon_name` vector or a `taxon` vector if we want the synonyms to be incorperated. 


## fg_name_search

```{r}
fg_name_search(q = "Gymnopus")
```

A lot of these columns could be made into a `taxon_name` vector or a `taxon` vector if we want the synonyms to be incorperated. 


## id2name

```{r}
id2name(19322, db = "itis")
```

Can return `taxon_name` vector

