
```{r echo = FALSE}
options(max.print = 100, width = 100)
```


```{r eval = FALSE}
# devtools::install_github("ropensci/taxa", ref = "vectorize")
```

```{r}
library(taxa)
```


I used the `vctrs` package to try out making S3 vector-like versions of classes to hold taxon databases and taxon id.
If this works out, I would like to do the same for taxon names, ranks, and taxa as a whole, which would include the ids, names, and ranks.
The ids, names, and ranks each can hold their own database ID.

## taxon_db 

The `taxon_db` class is just a character vector of database names with extra validity checks to make sure that names are valid.

```{r}
taxon_db()
taxon_db('ncbi')
taxon_db(rep('itis', 100))
taxon_db(rep('itis', 2000))
```

The acceptable databases are stored in `database_definitions` in the same way knitr chunk options are stored and changed.

```{r}
database_definitions$get()
```

All names in `taxon_db` objects must be in this list or errors are thrown:

```{r error=TRUE}
taxon_db(c('ncbi', 'itis', 'custom', 'tps'))
```

This also happens on assigning values:

```{r error=TRUE}
x = taxon_db(rep('itis', 10))
x
x[2:3] <- "ncbi"
x
x[2:3] <- "custom"
```

However, you can add new databases to the list:

```{r}
database_definitions$set(
  name = "custom",
  url = "http://www.my_tax_database.com",
  description = "I just made this up",
  id_regex = ".*"
)
```

And then you can use that name too:

```{r}
x[2:3] <- "custom"
x
```


## taxon_id

The `taxon_id` class stores taxon IDs and their associated database (as a `taxon_db` vector).

```{r}
taxon_id()
taxon_id(1:10)
taxon_id(1:10, db = 'ncbi')
taxon_id(1:10000, db = 'itis')
```

> Run code in an R console to see colored output. I forgot how to make it work in Rmarkdown.

Since a `taxon_db` object is used, database names must be valid:

```{r error=TRUE}
taxon_id(1:10, db = 'custom2')
```

The database values can be accessed and set with `taxon_db`:

```{r}
x <- taxon_id(1:10, db = 'ncbi')
x
taxon_db(x)
taxon_db(x) <- 'itis'
x
taxon_db(x)[1:3] <- 'tps'
x
```

... if they are valid:

```{r error=TRUE}
taxon_db(x)[1:3] <- 'custom2'
```

If a database is defined, then the ids must match the ID regex for that database in `database_definitions$get()`:

```{r error=TRUE}
taxon_id(letters[1:3], c('ncbi', 'itis', 'itis'))
taxon_id(letters[1:3])
```

You can set taxon id values with numbers or characters, but only `taxon_id` inputs can supply a defined database.

```{r error=TRUE}
x <- taxon_id(1:10, db = 'ncbi')
x
x[3:4] <- 1
x[6:7] <- c("a", "b")
x[8:9] <- taxon_id(1, db = c('itis', 'tps'))
x
```


## Use in tables

The `taxon_id` prints well in `data.frame`s and `tibble`s, using color when possible:

```{r}
data.frame(x = taxon_id(1:10, "ncbi"), y = taxon_db(rep("itis", 10)), z = 1:10)
tibble::tibble(x = taxon_id(1:10, "ncbi"), y = taxon_db(rep("itis", 10)), z = 1:10)
str(taxon_id(1:10, 'ncbi'))
```


## Plans for taxon name and rank

`taxon_name` and `taxon_rank` would be very similar to `taxon_id`.
`taxon_rank` would either be based on a factor, or be factor-like and would have database-based validity checks for rank names and order.


## Plans for taxon

`taxon` would contain at least `taxon_name`, `taxon_rank` and `taxon_id` vectors.
It would probably also have authority information.
It could also have field for arbitrary information, but this might not be needed since it could be in a table that could contain that info in other columns.


## Plans for taxonomy, taxmap, and hierarcy

Probably leave as R6 for now.


